\documentclass{article}
\usepackage{hyperref}

\title{A brief user's guide to Jedd}
\author{Ond\v{r}ej Lhot\'ak}

\begin{document}
\maketitle

\newcommand{\jeddpath}[1]{\href{http://svn.sable.mcgill.ca/soot/jedd/trunk/#1}{\nolinkurl{#1}}}

\section{Preliminaries}
This mini-tutorial assumes that the reader has
read the paper~\cite{Lhotak:2004:JBR} about Jedd that
was presented at PLDI 2004, and is available at
\url{http://www.sable.mcgill.ca/publications/papers/#pldi2004}.

\section{Example}
The Jedd distribution contains a directory called \jeddpath{examples}
containing sample Jedd code. Currently, it contains a single example,
\jeddpath{examples/pointsto}. This is a Jedd version of the BDD-based points-to
analysis from~\cite{Berndl:2003:PAU}.

\section{Jedd source files}
Source files to be processed by Jedd must have one of the extensions
{\tt .jedd} or {\tt .java}. It is customary to use the extension {\tt .jedd}
for files containing Jedd-specific constructs, and {\tt .java} for files
containing plain Java.

Jedd files should import the package {\tt jedd.*} from the Jedd runtime
library. This package contains interface classes with methods that can
be called by Jedd programs. In particular, the {\tt jedd.Jedd} class
is a singleton containing methods affecting the behaviour of Jedd in
general, and {\tt jedd.Relation} is an interface listing the methods
that can be called on any Jedd relation type. Jedd files should not
import the package {\tt jedd.internal.*}.

\section{Defining numberers, domains, attributes and physical domains}
The first step in writing a Jedd program is to define the numberers,
domains, attributes, and physical domains that will be used. A numberer
is a class that generates and maintains a mapping between objects and
non-negative integers, and must be implemented by the programmer. A
domain is a set of objects that will form the basis of Jedd relations.
Each domain must have an associated numberer for its objects. An
attribute is a domain with an assigned name, used to distinguish
multiple instances of a domain within the same relation. A physical
domain is a set of BDD bit positions to which Jedd may map an attribute
of a relation.

A numberer is a plain Java class implementing the {\tt jedd.Numberer}
interface. See the file \jeddpath{examples/pointsto/src/domains/IntegerNumberer.java},
which assigns to {\tt Integer} objects the integer that is their value.

A domain is similar to a Java class, but is defined with a
slightly different syntax: the class name is immediately followed
by an integer constant in parentheses. See the file 
\jeddpath{examples/pointsto/src/domains/Var.jedd} for an example.
The integer constant specifies how many bits are to be used to represent
the domain. The maximum number of objects in the domain is $2^b$, where $b$
is the number of bits specified. Each domain must extend the {\tt jedd.Domain}
class and implement the {\tt numberer()} method, which returns the
numberer for the domain.

An attribute is defined similarly to a domain, but the integer constant
number of bits is replaced with the name of the domain of the attribute.
See the file \jeddpath{examples/pointsto/src/attributes/var.jedd} for an
example with the domain {\tt Var}. An attribute must extend the 
{\tt jedd.Attribute} class. However, it should not implement its
abstract method {\tt domain()}; Jedd will implement it for you.

A physical domain is defined similarly to a domain or attribute, but
the parentheses following its name are empty. See the file 
\jeddpath{examples/pointsto/src/physical_domains/V1} for an example.
Each physical domain must extend the class {\tt jedd.PhysicalDomain}.

\section{Selecting a backend}
Jedd currently supports four different BDD libraries as backends:
BuDDy, CUDD, SableJBDD, and JavaBDD. BuDDy is the backend which has
the most complete support in Jedd, which is the most tested, and which
tends to perform best. BuDDy and CUDD are C libraries, so they require
that their shared library ({\tt .so} or {\tt .dll}) files be available
on the {\tt LD\_LIBRARY\_PATH}. Before using Jedd in you program,
you must select one of the backends by calling
{\tt jedd.Jedd.v().setBackend()}. The argument to this method
should be one of {\tt "buddy"}, {\tt "cudd"}, {\tt "sablejbdd"} or
{\tt "javabdd"}.

\section{Selecting a physical domain ordering (optional)}
By default, Jedd places the various physical domains one after the next
in the BDD. For performance reasons, you may want to select a different
ordering. This is done by calling {\tt jedd.Jedd.v().setOrder()}.
The first argument is an array containing the physical domains in the order
in which they should appear in the BDD, from the top of the BDD to the bottom
(constant nodes). To get the physical domain instance (to be placed in the array)
for a given physical domain, call the method {\tt v()} on it (see the example
in \jeddpath{examples/pointsto/src/Prop.jedd}). To interleave the bits of
several physical domains, place them in a separate array, and then
include this array as an element of the main array. The {\tt jedd.Jedd.setOrder()}
method takes a second, boolean argument that determines the order of bits within
each domain. Set it to {\tt true} to order the bits from most significant to least
significant when moving down the BDD from top to bottom, or to {\tt false} for the
opposite order.

\section{Writing Jedd code}
The Jedd grammar and explanations of its operators appear in~\cite{Lhotak:2004:JBR},
and are outside the scope of this guide. The paper also includes various
examples of Jedd code. Refer also to the points-to analysis example in
\jeddpath{examples/pointsto/src/Prop.jedd}.

The javadoc documentation of the (rather small) API available to Jedd programs
is available in \jeddpath{doc/api}. In particular, this includes the
{\tt jedd.Jedd} class with methods controlling the behaviour of Jedd in general,
and the {\tt jedd.Relation} interface of methods that can be called on any
relation type.

\section{Compiling Jedd code}
The Jedd compiler is invoked with the command {\tt java jedd.Main}.
It uses the same command-line format as Polyglot, with two additional
switches for specifying the path to a SAT solver ({\tt -s}) and a
SAT core extractor ({\tt -sc}). The simplest way to compile a project
is to list all the {\tt .jedd} files on the command line. This will
compile them to {\tt .java} files, and run {\tt javac} on them to
compile them to classfiles. The {\tt -c} switch disables the {\tt javac}
pass. If your project consists of both {\tt .jedd} and {\tt .java}
files, you can put them all on the command line, but be warned that
Polyglot will overwrite your {\tt .java} files unless you specify
an alternate output directory with the {\tt -d} switch.

The points-to analysis example includes a simple Ant build file
which can be modified for use in other projects.

\section{Using the profiler (optional)}
To use the profiler, it must be enabled before the computation to be profiled
begins by calling {\tt jedd.Jedd.v().enableProfiling()}. At the end of
the computation, the recorded profiling data can be written to a file
in SQL format by calling {\tt jedd.Jedd.v().outputProfile()} with a
{\tt java.io.PrintStream}. See the file \jeddpath{examples/pointsto/src/Prop.jedd}
for an example use of the profiler.

Viewing the profile data requires an SQL database and a CGI-capable web
server. The CGI scripts (found in the \jeddpath{profile_view} directory in the Jedd
distribution) are specific to SQLite, but should work with any
web server. They expect the profiling data in a database called
{\tt profile.db}, in the same directory as the scripts. This file can be
generated by piping the SQL file to SQLite with the command\\
{\tt cat profile.sql | sqlite profile.db}\\
(assuming the SQL file
is {\tt profile.sql}). thttpd can be started with the command:\\
{\tt /usr/sbin/thttpd -d /directory/with/cgi/scripts -p 8080 -c '*.cgi'}
(where {\tt /directory/with/cgi/scripts} is replaced with the directory
containing the Jedd CGI scripts from \jeddpath{profile_view}). This starts
the web server on port 8080. To view the profiling data, point your
web browser to\\
\url{http://127.0.0.1:8080/main.cgi}.

\bibliographystyle{alpha}
\bibliography{local}

\end{document}
